\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-03 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updated

%%%%%%%%%%%%%
\subsection{Comparison of flow pattern under the same simulation set-up with an additional inlet}

@@

\begin{figure}
\begin{center}
\subfigure[map of permeabilities]{\includegraphics[width=0.75\textwidth]{./Pics1/map_of_perm_1.pdf}}\\[2mm]%
\subfigure[flow at start]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_100.pdf}}%
\caption{In fig.(a) there is the map of permeabilities as this will be implemented in this case while in fig.(b) there is an illustration of the adaptive mesh just in the beggining of the simulation.}
\label{fig:3testcase_a}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_500.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_1500.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_3000.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_end.pdf}}%
\caption{from top to bottom, these are screenshots taken from the test-case of MR=$100$ using fixed and unstructured mesh for the same time steps, t=500(a), t=1500(b), t=3000(c), t=end(d)}
\label{fig:4testcase}
\end{center}
\end{figure}

\medskip
The first example is a simple comparison between two cases using fixed mesh and mobility ratios of $1$ and $10$ respectively. The aspect ratio of the $2D$ domain is 2 and it favors viscous fingering effects while temperature is not taken into account (assume isothermal environment) and gravity effects are negligible. Fluid $1$ is initially injected from the left hand side with velocity $u_{inj}=1$. The domain is fully saturated with fluid $2$. At the outlet (right hand side) boundary pressure levels are set to zero while no slip boundary conditions are implied at the upper and lower part of the formation. Fixed time step sizes of $1 \times 10^{-3}$ for both the fixed meshes and the CFL number for this set of simulations is $0.5$. The $P_{1}DGP_{2}$ element type with piecewise linear description of veleocity within each element and quadratic pressure representation as part of the OCVFEM method along with the discontinuous Galerkin method are implied and composing the mesh. Saturation is located at the pressure nodes. Five different regions of 5 distinct absolute permeabilities (K) are introduced (fig.\ref{fig:5regions_domain}). The red square region is saturated with a $50\%-50\%$ mixture of both phase $1$ and $2$.

Two general behaviors for the interface may occur: (i) If a more viscous fluid displaces a less viscous one or a less viscous displaces a more viscous with a low average velocity, the interface is flat; (ii) if a less viscous fluid displaces a more viscous one with intermediate values of the average velocity, single or mutliple fingers will be observed.

@@

%%%%%%%%%%%%%%%
\begin{comment}
\begin{figure}[h] 
\begin{center}
\subfigure[MR1]{%
\label{fig:}
\includegraphics[width=0.6\textwidth]{./Pics/MR1.pdf}
}%
\subfigure[Recovery for MR1]{%
\label{fig:}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR1_1.png}
}\\
\subfigure[MR10]{%
\label{fig:}
\includegraphics[width=0.6\textwidth]{./Pics/MR10.pdf}
}%
\subfigure[Recovery for MR10]{%
\label{fig:}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR10_10.pdf}
}\\
\subfigure[MR100]{%
\label{fig:}
\includegraphics[width=0.6\textwidth]{./Pics/MR100_100_start.pdf}
}%
\subfigure[Recovery for MR100]{%
\label{fig:}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR100_100.pdf}
}\\ 
\end{center}
\caption{Comparison of different MR's and their recovery rates. The comparison among the MR = $1$,MR = $10$ and MR = $100$ has been done using the boundary conditions and set upt as this has already been describe in the beginning of the section \ref{section:results}. }
\label{fig:recoveryrates}
\end{figure}
\end{comment}

@@

\medskip
A number of different cases have been illustrated in section \ref{section:results}. Starting from tested experimental cases and geomtrical domains to verify that the numerical model is working and set up some initial benchmarks, to more complicated cases where various scenarios were implied. Observations made on various mobility ratios, geometrical domains and permeability zones as well as on the role of interfacial forces on the developement of the flows. For immiscible displacements upon which only viscous and capillary forces are acting (being horizontal flow, gravity can be ignored), the expected displacement patterns are either stable front displacement or unstable front displacement. In the latter case the instability arises due to either capillary or viscous fingering depending on capillary number ($N_{c}$) but mainly on mobility ratio (MR). If the mobility ratio is significantly less than $5$ (MR=$3$ has been used) , it is expected to find capillary fingering at relatively low flow rate and viscous fingering at relatively high flow rate. On the other hand, when the viscosity ratio is greater than $5$, the two basic forms of immiscible displacement are capillary fingering at relatively low flow rate and stable displacement at relatively high flow rate. Therefore for an i.e., water/oil system, the first condition applies as the viscosity of injected water is less than the viscosity of displaced oil. Hence, the displacement form could be either capillary fingering at relatively low injection rates or viscous fingering at higher injection rates.

\medskip
We have numerically computed the evolution of the initial condition and for various values of the viscosity contrast under the same initial conditions and surface tension. Even though the commercial simulators provide the user with option of an unstructured type of mesh still they are not able to dynamically adapt the mesh as the simulation develops and they increase the computational cost as the saturation front changes slowly. In this paper the use and benefits of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows was introduced.

@@

\KCnote{Having confirmed that equal viscosity fluids do not finger into each other, a set of experiments were completed at different mobility ratios. Keeping both the thickness and the injection rate constant, it was observed that increasing the mobility ratio decreases the finger width, which subsequently leads to more fingers and subsequently more complexity at a given distance away from the injection point. Tip-splitting, a phenomenon in which the initial stage of fingers split into two or more new fingers, can also be observed at higher mobility ratios.}

>>>>

%%%%%%%%%%%%%
\subsection{Comparison of flow pattern with a different simulation set-up with and an additional inlet region.}

@@

\begin{figure}
\centering
\subfigure[map of permeabilities]{\includegraphics[width=0.75\textwidth]{./Pics1/map_of_perm_1.pdf}}\\[2mm]%
\subfigure[flow at start]{\includegraphics[width=0.75\textwidth]{./Pics1/5reg_dinlet_adapt_100.pdf}}%
\caption{In fig.(a) there is the map of permeabilities as this will be implemented in this case while in fig.(b) there is an illustration of the adaptive mesh just in the beggining of the simulation.}
\label{fig:3testcase_a}
\end{figure}

@@

\begin{figure}
\centering
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_500.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_1500.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_3000.pdf}}\\[2mm]%
\subfigure[]{\includegraphics[width=0.5\textwidth]{./Pics1/mr100_fixed/mr100_fixed_end.pdf}}%
\caption{from top to bottom, these are screenshots taken from the test-case of MR=$100$ using fixed and unstructured mesh for the same time steps, t=500(a), t=1500(b), t=3000(c), t=end(d)}
\label{fig:4testcase}
\end{figure}

\medskip
The first example is a simple comparison between two cases using fixed mesh and mobility ratios of $1$ and $10$ respectively. 
%The aspect ratio of the $2D$ domain is 2 and it favors viscous fingering effects while temperature is not taken into account (assume isothermal environment) and gravity effects are negligible. 
Fluid $1$ is initially injected from the left hand side with velocity $u_{inj}=1$. The domain is fully saturated with fluid $2$. At the outlet (right hand side) boundary pressure levels are set to zero while no slip boundary conditions are implied at the upper and lower part of the formation. Fixed time step sizes of $1 \times 10^{-3}$ for both the fixed meshes and the CFL number for this set of simulations is $0.5$. The $P_{1}DGP_{2}$ element type with piecewise linear description of veleocity within each element and quadratic pressure representation as part of the OCVFEM method along with the discontinuous Galerkin method are implied and composing the mesh. Saturation is located at the pressure nodes. Five different regions of 5 distinct absolute permeabilities (K) are introduced (fig.\ref{fig:5regions_domain}). The red square region is saturated with a $50\%-50\%$ mixture of both phase $1$ and $2$.

Two general behaviors for the interface may occur: (i) If a more viscous fluid displaces a less viscous one or a less viscous displaces a more viscous with a low average velocity, the interface is flat; (ii) if a less viscous fluid displaces a more viscous one with intermediate values of the average velocity, single or multiple fingers will be observed.

@@

@@

\medskip
A number of different cases have been illustrated in Section \ref{section:results}. Starting from tested experimental cases and geometrical domains to verify that the numerical model is working and set up some initial benchmarks, to more complicated cases where various scenarios were implied. Observations made on various mobility ratios, geometrical domains and permeability zones as well as on the role of interfacial forces on the developement of the flows. For immiscible displacements upon which only viscous and capillary forces are acting (being horizontal flow, gravity can be ignored), the expected displacement patterns are either stable front displacement or unstable front displacement. In the latter case the instability arises due to either capillary or viscous fingering depending on capillary number ($N_{c}$) but mainly on mobility ratio (MR). If the mobility ratio is significantly less than $5$ (MR=$3$ has been used) , it is expected to find capillary fingering at relatively low flow rate and viscous fingering at relatively high flow rate. On the other hand, when the viscosity ratio is greater than $5$, the two basic forms of immiscible displacement are capillary fingering at relatively low flow rate and stable displacement at relatively high flow rate. Therefore for an i.e., water/oil system, the first condition applies as the viscosity of injected water is less than the viscosity of displaced oil. Hence, the displacement form could be either capillary fingering at relatively low injection rates or viscous fingering at higher injection rates.

%\medskip
%We have numerically computed the evolution of the initial condition and for various values of the viscosity contrast under the same initial conditions and surface tension. Even though the commercial simulators provide the user with option of an unstructured type of mesh still they are not able to dynamically adapt the mesh as the simulation develops and they increase the computational cost as the saturation front changes slowly. In this paper the use and benefits of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows was introduced.

@@

%\KCnote{Having confirmed that equal viscosity fluids do not finger into each other, a set of experiments were completed at different mobility ratios. Keeping both the thickness and the injection rate constant, it was observed that increasing the mobility ratio decreases the finger width, which subsequently leads to more fingers and subsequently more complexity at a given distance away from the injection point. Tip-splitting, a phenomenon in which the initial stage of fingers split into two or more new fingers, can also be observed at higher mobility ratios.}
